Exercise 2: Quality Attributes and Architectural Design of the HS07 System

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
Updated sectioning

\subsection{Selection}
Here we put the scenarios to the vote. This is done by emulation of a real-life situation where the actual stakeholder would
be part of the process. We take on the roles of the stakeholders in their absence, and chose which scenarios would be more
interesting for us.
\subsection{Stakeholders and interests}
\begin{itemize}
\item User (homeowner)
\begin{itemize}
\item Easy to use and configure
\item Always available
\item Safe (will not make an error and burn down the house)
\item Secure (other people cannot control my home)
\end{itemize}
\item Producer
\begin{itemize}
\item Cheap
\item Production line (modifiability for other uses)
\end{itemize}
\item Technical installer
\begin{itemize}
\item Easily deployable (installable)
\end{itemize}
\item Developer
\begin{itemize}
\item Easily adjustable to new requirements
\item Scalable
\item Remote update (upgrading online)
\item Remote maintenance (debugging online)
\item Automatically testable (regression test)
\item Developing in parallel
\end{itemize}
\item Insurance company
\begin{itemize}
\item Safe with respect to fires (radiators) and break ins (door), ...
\end{itemize}
\end{itemize}
\subsection{Identical scenarios}
Due to the distributed location of the participants in the brainstorming multiple "similar" scenarios was created. In order to avoid having the same scenario represented twice with a slightly different wording, the below list show the scenarios which will be considered identical, meaning that their votes will be accumulated.
\begin{itemize}
\item 2 and 7
\item 6 and 16
\item 10 and 17
\item 9 and 11 and 18
\item 1 and 20
\item 5 and 21
\end{itemize}
\subsection{The Vote}
\subsection{First vote}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Person & Stakeholder Role & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
\hline
Jesper&User&4&5&6/16&20&21&2/7\\
Lars&Developer&19&19&5/21&9/11/18&9/11/18&13\\
Anders&Producer&2/7&9/11/18&9/11/18&12&10/17&10/17\\
\hline
\end{tabular}
\subsection{Second vote}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Person & Stakeholder Role & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
\hline
Jesper&Developer&9/11/18&9/11/18&9/11/18&19&13&6/16\\
Lars&User&1/20&4&6/16&15&4&5/21\\
Anders&Producer&9/11/18&6/16&5/21&5/21&12&10/17\\
\hline
\end{tabular}
\subsection{Voting Results}
This is the result of the vote in a prioritized order.
\begin{enumerate}
\item 9/11/18 (8 votes)
\item 5/21 (6 votes)
\item 6/16 (4 votes)
\item 19 (3 votes)
\item 4 (3 votes)
\item 13 (2 votes)
\item 1/20 (2 votes)
\item 2/7 (2 votes)
\item 10/17 (3 votes)
\item 12 (2 votes)
\item 15 (1 votes)
\item 3 (0 votes)
\item 8 (0 votes)
\item 14 (0 votes)
\end{enumerate}
\subsection{Refinement}
The process of refining the chosen scenarios centers around specification of various elements of the scenario. Stimulus, Response, Response Measurement.

@@

\subsection{Adding new sensor/actuator}
Localized to changes within the given sensor or actuator type. Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case.

\subsection{Adding different types}
Update by addition will need to be built into the architecture in the sensor and actuator subsystems to encompass new types apart from thermometers and radiators. Radiators and thermometers are hardcoded into the gateway, it will need to be more abstract.

@@

\subsection{Refined Scenario \#1}
\subsection{Tactics}
\subsection{Modifiablity}
\begin{itemize}
\item New Thermometer/Radiator implementation
\begin{itemize}
\item In the present implementation the ThemometerService has a hard-coded reference to Themometer, and having a new operation in the implementation is bad design if modifiability is desired, as it promoted change by modification. Changing the ThemometerService to use e.g. AbstractFactory to create the given Themometer instance will allow a new Themometer to be implemented and deployed by merely adding a class and a factory, which is modification by addition and then writing a main for the device, which uses this factory and the generic ThermometerService. At present the getController returns an object and the AbstractFactory may be implemented to also return this. It would however be an advantage to define a Thermometer interface in order to ensure that the implementations meets the protocol (has getTemperature method). This imply several tactics, as it \emph{localize modifications} to the temperature implementation. If a Temperature interface is created, then this uses \emph{anticipate expected changes} as the interface must contain the methods, which is expected to be needed, not just now but in the future. \emph{Generalize the module} is used in the fact that Objects are exchanged between Gateway and sensor/actuator, giving the Gateway to work on any instance of a class. 
\item The use of HTTP between Gatwway, sensor and actuator removes the syntactical dependency replacing it with runtime checking (by the Invoker), with the advantages and disadvantages it has. There is also a strong \emph{Location dependency} from all sensore/actuators to the gateway, as they must know where it is located (no discovery). It may be updated with a DNS-lookup to make this weaker.
\item As the Gateway must support all sensors/actuators in its run algorithm there is a \emph{QoS dependency} that this must not take longer than the period by which the run-method is expected to iterate.
\item The Gateway already support \emph{defer binding time} as the actuator/sensor registers with the Gateway at runtime.
\end{itemize}
\item New sensor/actuator type
\begin{itemize}
\item The current implementation of the Gateway supports \emph{defer binding time}, yet as the Gateway run method has a hardcoded dependency on the semantics of the radiator and thermometer (through the getTemperature and notifyObserver command syntax) it is not possible to simply register other types. The Gateway would have to use \emph{Generalize the module} in order to allow more general actuators and sensors to register, as opposed to only those that have radiator and thermometer semantics.
\end{itemize}
\end{itemize}
\subsection{Testability}
\subsection{Refined Scenario \#3}
Scenario \#3 requires the HS07 system to provide a web interface that can be used to change the temperature in the house. The response requirements sets some requirements to the availability of the http service.
\subsection{Tactics}
\subsection{Availability}
As the user may experience non-responsiveness 0,1\% of the time it is believed that this can be done with a single hardware unit (no HW-reduadance, e.g. \emph{voting}). At the same time the requirement of 30 seconds 1\% of the time is sufficient time to restart the web-service, and it is therefore not required to have \emph{active redundancy}. Finally it is believed that an availability of 5 seconds 98,9\% of the times, require a form of redundancy, and for this reason \emph{passive redundancy} is chosen. As an example of a design may be mentioned the Apache web-server design, which is redundant processes. A number of processes is spawned to handle HTTP/GET requests, and it uses some kind of fault detection scheme to detect if one become non-responsive. If this happens it is simply killed and an identical process is spawned in its place.

\subsection{Usability}
A web interface must be provided to support the requirement that the user can change the temperature from a web page. In order to maintain the modifiability of the system, a \emph{Model-View-Controller} pattern will be used to separate the web interface from the gateway. A web backend process will be added to limit the web access to the model thereby hiding irrelevant information. The model needs to maintain all information about the system that is relevant to the user - \emph{Maintain a model of the system}.

@@

\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/mv-model",scale=0.8 }}
\caption{Module View of the web interface in the redesigned HS07 system}
\label{fig:mv-model}
\end{figure}

\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/mv-factory",scale=0.8 }}
\caption{Module View of the abstract factory in the redesigned HS07 system}
\label{fig:mv-factory}
\end{figure}

@@

\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/Active object",scale=0.8 }}
\caption{Active Objects in the redesigned HS07 system}
\label{fig:activeobjects}
\end{figure}

@@

\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/seq-stt",scale=0.8 }}
\caption{Sequence Diagram for the web requests in the redesigned HS07 system}
\label{fig:seq-stt}
\end{figure}

>>>>

\subsection{Selection}
Here we put the scenarios to the vote. This is done by emulation of a real-life situation where the actual stakeholder would
be part of the process. We take on the roles of the stakeholders in their absence, and chose which scenarios would be more
interesting for us.
\subsection{Stakeholders and interests}
\begin{itemize}
\item User (homeowner)
\begin{itemize}
\item Easy to use and configure
\item Always available
\item Safe (will not make an error and burn down the house)
\item Secure (other people cannot control my home)
\end{itemize}
\item Producer
\begin{itemize}
\item Cheap
\item Production line (modifiability for other uses)
\end{itemize}
\item Technical installer
\begin{itemize}
\item Easily deployable (installable)
\end{itemize}
\item Developer
\begin{itemize}
\item Easily adjustable to new requirements
\item Scalable
\item Remote update (upgrading online)
\item Remote maintenance (debugging online)
\item Automatically testable (regression test)
\item Developing in parallel
\end{itemize}
\item Insurance company
\begin{itemize}
\item Safe with respect to fires (radiators) and break ins (door), ...
\end{itemize}
\end{itemize}
\subsection{Identical scenarios}
Due to the distributed location of the participants in the brainstorming multiple "similar" scenarios was created. In order to avoid having the same scenario represented twice with a slightly different wording, the below list show the scenarios which will be considered identical, meaning that their votes will be accumulated.
\begin{itemize}
\item 2 and 7
\item 6 and 16
\item 10 and 17
\item 9 and 11 and 18
\item 1 and 20
\item 5 and 21
\end{itemize}
\subsection{First vote}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Person & Stakeholder Role & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
\hline
Jesper&User&4&5&6/16&20&21&2/7\\
Lars&Developer&19&19&5/21&9/11/18&9/11/18&13\\
Anders&Producer&2/7&9/11/18&9/11/18&12&10/17&10/17\\
\hline
\end{tabular}
\subsection{Second vote}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Person & Stakeholder Role & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
\hline
Jesper&Developer&9/11/18&9/11/18&9/11/18&19&13&6/16\\
Lars&User&1/20&4&6/16&15&4&5/21\\
Anders&Producer&9/11/18&6/16&5/21&5/21&12&10/17\\
\hline
\end{tabular}
\subsection{Voting Results}
This is the result of the vote in a prioritized order.
\begin{enumerate}
\item 9/11/18 (8 votes)
\item 5/21 (6 votes)
\item 6/16 (4 votes)
\item 19 (3 votes)
\item 4 (3 votes)
\item 13 (2 votes)
\item 1/20 (2 votes)
\item 2/7 (2 votes)
\item 10/17 (3 votes)
\item 12 (2 votes)
\item 15 (1 votes)
\item 3 (0 votes)
\item 8 (0 votes)
\item 14 (0 votes)
\end{enumerate}
\subsection{Refinement}
The process of refining the chosen scenarios centers around specification of various elements of the scenario. Stimulus, Response, Response Measurement.

@@

\subsubsection{Adding new sensor/actuator}
Localized to changes within the given sensor or actuator type. Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case.

\subsubsection{Adding different types}
Update by addition will need to be built into the architecture in the sensor and actuator subsystems to encompass new types apart from thermometers and radiators. Radiators and thermometers are hardcoded into the gateway, it will need to be more abstract.

@@

\subsection{Refined Scenario \#1 - Tactics}
\subsubsection{Modifiablity}
\begin{itemize}
\item New Thermometer/Radiator implementation
\begin{itemize}
\item In the present implementation the ThemometerService has a hard-coded reference to Themometer, and having a new operation in the implementation is bad design if modifiability is desired, as it promoted change by modification. Changing the ThemometerService to use e.g. AbstractFactory to create the given Themometer instance will allow a new Themometer to be implemented and deployed by merely adding a class and a factory, which is modification by addition and then writing a main for the device, which uses this factory and the generic ThermometerService. At present the getController returns an object and the AbstractFactory may be implemented to also return this. It would however be an advantage to define a Thermometer interface in order to ensure that the implementations meets the protocol (has getTemperature method). This imply several tactics, as it \emph{localize modifications} to the temperature implementation. If a Temperature interface is created, then this uses \emph{anticipate expected changes} as the interface must contain the methods, which is expected to be needed, not just now but in the future. \emph{Generalize the module} is used in the fact that Objects are exchanged between Gateway and sensor/actuator, giving the Gateway to work on any instance of a class. 
\item The use of HTTP between Gatwway, sensor and actuator removes the syntactical dependency replacing it with runtime checking (by the Invoker), with the advantages and disadvantages it has. There is also a strong \emph{Location dependency} from all sensore/actuators to the gateway, as they must know where it is located (no discovery). It may be updated with a DNS-lookup to make this weaker.
\item As the Gateway must support all sensors/actuators in its run algorithm there is a \emph{QoS dependency} that this must not take longer than the period by which the run-method is expected to iterate.
\item The Gateway already support \emph{defer binding time} as the actuator/sensor registers with the Gateway at runtime.
\end{itemize}
\item New sensor/actuator type
\begin{itemize}
\item The current implementation of the Gateway supports \emph{defer binding time}, yet as the Gateway run method has a hardcoded dependency on the semantics of the radiator and thermometer (through the getTemperature and notifyObserver command syntax) it is not possible to simply register other types. The Gateway would have to use \emph{Generalize the module} in order to allow more general actuators and sensors to register, as opposed to only those that have radiator and thermometer semantics.
\end{itemize}
\end{itemize}
\subsubsection{Testability}
Tactics for testability include \emph{Test harness} which is essentially a needed setup for regression testing. The test harness for the HS07 system will consist of a series of JUnit test cases on a modular level. On a larger scale, the deployment on a PC, shown in /cite{SAiP1H1} is one big test scenario where sensors and actuators a stubbed using visual inspection for verification. This harness will continue to be the main test setup for introduction into the gateway/model parts of the system.
\subsection{Refined Scenario \#3 - Tactics}
Scenario \#3 requires the HS07 system to provide a web interface that can be used to change the temperature in the house. The response requirements sets some requirements to the availability of the http service.
\subsubsection{Availability}
As the user may experience non-responsiveness 0,1\% of the time it is believed that this can be done with a single hardware unit (no HW-reduadance, e.g. \emph{voting}). At the same time the requirement of 30 seconds 1\% of the time is sufficient time to restart the web-service, and it is therefore not required to have \emph{active redundancy}. Finally it is believed that an availability of 5 seconds 98,9\% of the times, require a form of redundancy, and for this reason \emph{passive redundancy} is chosen. As an example of a design may be mentioned the Apache web-server design, which is redundant processes. A number of processes is spawned to handle HTTP/GET requests, and it uses some kind of fault detection scheme to detect if one become non-responsive. If this happens it is simply killed and an identical process is spawned in its place.

\subsubsection{Usability}
A web interface must be provided to support the requirement that the user can change the temperature from a web page. In order to maintain the modifiability of the system, a \emph{Model-View-Controller} pattern will be used to separate the web interface from the gateway. A web backend process will be added to limit the web access to the model thereby hiding irrelevant information. The model needs to maintain all information about the system that is relevant to the user - \emph{Maintain a model of the system}.

@@

\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/mv-model",scale=0.8 }}
\caption{Module View of the web interface in the redesigned HS07 system}
\label{fig:mv-model}
\end{figure}
Figure~\ref{fig:mv-model} illustrates the new packages needed in HS07 to include a web interface for the user. The model package creates a service with witch to get and set the values  need by the user to make a decision an any new target temperature. Both the gateway and the webapp uses their invokers to contact the model service for parameter manipulation.

The separation between the gateway and the model packages creates an opportunity for increase of security when that becomes a driving quality attribute, since the gateway information can be hidden within the system, letting the ModelService become the only exposed part.
\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/mv-factory",scale=0.8 }}
\caption{Module View of the abstract factory in the redesigned HS07 system}
\label{fig:mv-factory}
\end{figure}

@@

\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/Active object",scale=0.8 }}
\caption{Active Objects in the redesigned HS07 system}
\label{fig:activeobjects}
\end{figure}
Figure~\ref{fig:activeobjects} shows all the components of the HS07 system. The new parts are the ModelService and the HTTP Server, needed to create a suitable web interface for the user.
\clearpage

@@

\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/seq-stt",scale=0.8 }}
\caption{Sequence Diagram for the web requests in the redesigned HS07 system}
\label{fig:seq-stt}
\end{figure}
Figure~\ref{fig:seq-stt} shows the flow of contol when a web application uses the methods exposed throgh the model service. The amount of traffic on the model service will have no direct impact on the performance of the gateway service.

The gateway uses the same http interface as the webapp to set the value of the current average room temperature, and to read the values of the target temperature requested by the user.

If the ModelService is non responsive, it is important for the gateway service to continue to work, "guessing" the target temperature either by cache or by default values.
